The interactions of rational, pragmatic agents  \n lead to efficient language structure and use

### 2019-01-31 paper/paper.Rmd
updates Thurs morning.

Zipf articulated what is now considered a *functionalist* approach to language science -- analyzing language structure and use in terms of efficiency. Such an approach might reframe our opening question as follows: how does having property \textit{x} make using language $\ell$ more or less useful for communication? Among others, this efficiency-based framing has produced a rich set of theoretical and empirical targets exploring semantic typology [@RegierKempKay2015a], properties such as ambiguity [@Piantadosi2011a] and compositionality [@KirbyGriffithsSmith2014a], and the efficient use of reduction and redundancy in production [@LevyJaeger2007a; @GenzelCharniak2002a]. These projects typically employ information-theoretic principles to characterize efficiency in natural or simulated language corpora. We adopt a different approach in this work, deriving a measure of efficiency from first principles of \textit{efficient language use} and then showing how this can lead to communicatively efficient systems.\par

Functionalist theories commonly frame language efficiency in terms of a fundamental effort-assymetry underlying everyday communication: what is “hard” for a speaker is likely different than what is “hard” for a listener. Zipf gave the following intution for how this idea -- purely from the standpoint of speaker effort, an optimal language $\ell_{speaker}^*$ would tend toward a vocabulary of a single, low-cost word. Given such a language, the full set of potential meanings would be conveyed using only that word, i.e. $\ell_{speaker}^*$ would be fully ambiguous and all possible meanings would need to be disambiguated by a listener. From the standpoint of listener effort, an optimal language $\ell_{listener}^*$ would map all possible meanings to distinct words, removing a listener's need to disambiguate. In this example, speaker effort is related to *production cost* and listener effort to *understanding or disambiguation cost*. Clearly, natural languages fall between the two extremes of $\ell_{speaker}^*$ and $\ell_{listener}^*$. Zipf proposed that the particular lexicon-level properties he observed were a result of these competing forces -- the pressure to jointly minimize speaker and listener effort.\par

@@

In the current work we provide a case-study for the framing presented above. We choose a property of languages that could, in principle, vary freely, but shows strong regularities across languages. The explanadum is why this regularity holds. We examine ambiguity as our property, extending ideas by Piantadosi et al. [-@Piantadosi2011a]. We define a novel measure of efficiency that depends on the interactional behavior of speaker and listener agents. We adopt the reference game as our primary unit of interaction and model language users using the Rational Speech Act (RSA) framework -- a computional model of language use, which is nicely attested by experimental data on interaction. Using these basic ingredients, we show that the property of interest (ambiguity) is prevalent in languages that optimize our measure of efficiency. Further, we show how ambiguity is \textit{used} efficiently during local, in-the-moment interactions. The contributions of this work are twofold -- we derive a novel objective for measuring linguistic efficiency and also show how the reference game framework, in combination with formal models of communication, can be used to connect ideas about system-level efficiencies to in-the-moment language use.\par

@@

### Reference games
Zipf’s example of optimal speaker- and listener-languages took the form of a reference game. We adopt that formulation here, assuming these communication games as our basic unit of analysis. In this framework, speakers and listeners are aware of a set of objects $M$ (*meanings*) and are knowledgeable about the set of possible signals $U$ (*utterances*) that can be used to refer to a given meaning (see Figure 1). Utterances may have different relative costs, operationalized via a prior over utterances $P(U)$. Similarly, meanings differ in the relative degree to which they need to be talked about, operationalized as a prior over meanings $P(M)$\footnote{The prior over meanings are analogous to the \textit{need probabilities} assumed in previous work (Regier, Kemp \& Kay (2015).}. We consider a set of contexts $c\in C$ with an associated prior $P(C)$. Each context $c_i$ describes a different distribution over meanings e.g. $p(M|C=c_i) \neq p(M|C=c_j)$. Finally, we consider a set of communicative events $e \in E$ where $<u, m, c> = e$ is an utterance, meaning and context triple.\par

@@

### Speakers and listeners
The Rational Speech Act framework (RSA) is computational-level theory of pragmatic language use, which has been well attested in experimental work examining both speaker and listener communication behavior [@FrankGoodman2012a; @GoodmanFrank2016a]. RSA is a formalization of essential Gricean pragmatic principles -- agents reason about one another and their shared context [@Grice1975a]. We adopt RSA as our representational framework to model Gricean (rational and pragmatic) speaker-listeners in the reference game setting\footnote{See SI section 1 for a more detailed description of basic RSA speaker-listener defintions and to Goodman \& Frank [-@GoodmanFrank2016a] for an overview of its applications.}.\par

@@

\begin{equation}
\begin{split}
  = \mathbb{E}_{c \sim P(C)}[H_{cross}(P_{speaker}, P_{listener} | c; \ell)]\\
\end{split}
\end{equation}
From an information-theoretic perspective this objective is intuitive: The Cross-Entropy (CE) gives us a measure of dissimilarity between two distributions -- the average number of bits required to communicate under one distribution, given that the “true” distribution differs. In our case, we have an expectation over this term -- the expected difference between the distributions assumed by the speaker $P_{speaker}$ and listener $P_{listener}$ given a set of contexts $C$. In other words, an "efficient" language $\ell$ minimizes the distance between what speakers and listeners think.

@@

The task of understanding language is marked by a frequent need to handle various forms of ambiguity: lexical, syntactic, among others[@WasowPerforsBeaver2005a]. The ubiquity of this property, however, has been argued to provide evidence that langauges are not been optimized for communication\footnote{Chomsky's (2002) claim that "... language is not well designed, because you have such properties as ambiguity" is analogous to the Zipfian optimal Listener language $\ell_{listener}^*$.} [@Chomsky2002a].

@@

We re-iterate their argument here. Say we have two objects ($m_1$ and $m_2$), two utterances ($u_1$ and $u_2$), which differ in cost, and two languages ($\ell_1$ and $\ell_2$) which describe different utterance-meaning mappings. In language $\ell_1$, the low-cost $u_1$ can be used to refer to both $m_1$ and $m_2$ ($[[u_1]]_{\ell_1} = \{m_1, m_2 \}$), but the high-cost $u_2$ cannot be used at all ($[[u_2]]_{\ell_1} = \emptyset$). By contrast, in language $\ell_2$, $u_1$ can only refer to $m_1$ and $u_2$ can only refer to $m_2$ ($[[u_1]]_{\ell_2} = \{m_1\}$ and $[[u_2]]_{\ell_2} = \{m_2 \}$). While it is cheaper for a speaker to use $\ell_1$ (because speaking is always low-cost), it is more difficult for a listener (because $u_1$ is ambiguous). Crucially, however, if context is disambiguating then the speaker can use $u_1$ to refer to either $m_1$ or $m_2$ and $\ell_1$ should be preferred to $\ell_2$.\par

In the following experiments we explore two aspects of Piantadosi’s claim. First, we examine the efficient language *structure* aspect of their claim, exploring when the optimal linguistic system $\ell^*$ is most likely to contain ambiguous lexical items. In the second experiment, we explore an efficient language *use* aspect of the claim -- at what point in a conversation is it useful for a speaker to use ambiguous lexical material?\par

@@

We adopt the hypothesis presented by Piantadosi, formalizing their argument in our current framework -- showing that ambiguity is an efficient property under our CE objective in the reference game setting. We proceed by generating languages with different amounts of contextual support (varying the size of $|C|$). We search the space of languages, examining whether ones which minimize our objective contain ambiguity. If context leads to more efficient commnication, then optimal languages under our objective should be more likley to be ambiguous as the amount of context increases.\par

@@

We conduct $N=2000$ simulations. For each simulation we enumerate the set of *valid* languages in which $|U|=|M|=4$ (recall that $U$ is our set of utterances and $M$ our set of meanings). Note that a language $\ell \in L$ is "valid" so long as each possible meaning in $m \in M$ can be referred to by at least one form $u \in U$ (every column of $\ell$ has some non-zero assignment) and each form maps to at least one meaning (every column has some non-zero assignment). For a given simulation the goal is to find the language $\ell^*$ which minimizes our CE objective and then check to see if that language contains ambiguity.\par

@@

Following the argument given by Piantadosi et al. [-@Piantadosi2011a], we want to assess the impact of *context* on the presence of ambiguity in optimal languages. To do so we consider four conditions with $n=500$ simulations each (that is, 500 unique sets of $\{P(U), P(M|C), P(C)\}$. Our first is a *one-context* condition ($|C|=1$) -- only a single distribution over meanings $P(M)$. In our two-context condition ($|C| = 2$), we consider efficiency under both $P(M|C=c_1)$ as well as $P(M|C=c_2)$. Three- and four-context conditions corresponding accordingly.\par

@@

Figure 1. plots the proportion of optimal languages under each objective as a function of context size. The red line shows that as the number of contexts increases, so does the probability that an optimal language $\ell^*_{cross}$ contains ambiguity under our CE objective. For comparison we also plot the proportion of speaker-optimal $\ell^*_{speaker}$ (blue line) and listener-optimal $\ell^*_{listener}$ (green line) languages that contain ambiguity. In line with Zipf's predictions, if languages are designed only to minimize speaker effort then optimal languages always contain ambiguity. If languages are designed to minimize listener effort then ambiguity is always avoided.\par

In simulation 1 we explored the degree to which ambiguity was an efficient property when communication was contextualized. We showed that as contextual information increased, the optimal languages under our CE objective were more likely to have ambiguous lexical items. Importantly, we did not see this effect under alternative objectives which only emphasized speaker- or listener effort alone.\par

While our results indicate that ambiguity is an efficient property of contextualized language, these simulations assumed that agents could always disambiguate items in context. That is, in our four conditions both agents had perfect knowledge of the relevant conditional distributions ($P(M|C)$). This assumption may be too strong for describing much of day-to-day communication -- we seldom interact with others with perfect knowledge of the current context (or topic) of conversation. To explore how ambiguity may be \textit{used} efficiencly, we next explore a case in which the listener has imperfect knowledge of context at the start of the conversation, but may infer it from a discourse history in Simulation 2.\par

@@

We shift focus in Simulation 2 -- instead of examining the landscape of a space of languages under our CE objective, we consider a single language containing ambiguity, examining how that property is used over a discourse. If context is not informative with respect to meaning early in a discourse, but is at later positions, then an efficient speaking strategy would avoid ambiguous material early, saving it for when contextual information is disambiguating. In this toy example we can consider “context” as analogous to a “topic” of conversation. As an example, imagine a scenario in which a reader is beginning a magazine article. While they may have some knowledge about the article’s topic (perhaps from the title), they may not have complete knowledge of its contents, including the persons or events involved. In this setting, using an potential low-cost, but ambiguous pronoun early may lead to misunderstanding if context is sufficiently informative. But, if by a later position enough contextual information has accumulated, it may be efficient for the writer to use the ambiguous material.\par

@@

Over the course of a discourse $D$, the listener tries to infer both the current topic, $c_{current}$, as well as the particular meaning $m$ of a given utterance $u$. That is, we consider updated listener and speakers ($L(m, c|u,D;\ell)$, $S(u|m,c,D;\ell)$) in which both agents can track the history of previous utterances $D$. Importantly, an agent can attempt to infer the current topic of conversation $c_{current}$ using the discourse history $D$ (see SI section 3 for more detail).\par

@@

## Speaker agents
We consider three types of speaker models. Our \textit{Full pragmatics} agent, models a speaker who reasons about her listener and also has complete recall of the set of utterances in the discourse $D$. This speaker believes that the listener may not know the current topic $c_{current}$ at the start of the discrouse, but should infer it over the discourse. We compare two baselines models. The first, a \textit{Partial pragmatics} baseline describes a speaker who reasons about a listener, but assumes they have no access to discourse history. The second, a \textit{No pragmatics} speaker who does not consider a listener at all, but produces utterances according to the underlying langauge semantics ($\ell$) and topic probabilities ($p(M|C=c_{current}$) (see SI section 3 for details).\par

@@

Figure 2, (A) shows the empirical probability that a speaker uses an ambiguous utterance as a function of discourse position. The \textit{No pragmatic} baseline uses ambiguous utterances frequently and at a constant rate over the discourse and the \textit{Partial pragmatic} baseline avoids ambiguous utterances entirely. But, the \textit{Full pragmatic} model avoids ambiguous material only at the start of the discourse, employing it increasingly as the discourse proceeds. (C) tracks our CE objective for each model over the discourse. Note that the objective decreases for all three models, primarily driven by the listener updating his belief about the actual topic ($P(C=c_{current}|D)$). However, the objective declines more quickly under the \textit{Full} and \textit{Partial} pragmatic speakers as listener agents are better able to infer the correct context. Additionally, the difference in CE between the \textit{Full}- and \textit{Partial pragmatic} models at the end of the discourse is driven by the reduction in speaker costs described in (B). While speaker effort remains constant in both \textit{No pragmatic} and \textit{Partial pragmatic} baselines, effort declines in the \text{Full Pragmatic} model as she increasingly relies on ambiguous material later in the discourse.\par

>>>>

Zipf articulated what is now considered a *functionalist* approach to language science -- analyzing language structure and use in terms of efficiency. Such an approach might reframe our opening question as follows: how does having property \textit{x} make using language $\ell$ more or less useful for communication? Among others, this efficiency-based framing has produced a rich set of theoretical and empirical targets exploring semantic typology [@RegierKempKay2015a], properties such as ambiguity [@Piantadosi2011a] and compositionality [@KirbyGriffithsSmith2014a], and the efficient use of reduction and redundancy in production [@LevyJaeger2007a; @GenzelCharniak2002a]. These projects typically employ information-theoretic principles to characterize efficiency in natural or simulated language corpora. We adopt a different approach in this work, deriving a measure of efficiency from first principles of \textit{efficient language use} and show how this can lead to communicatively efficient systems.\par

Functionalist theories commonly frame language efficiency in terms of a fundamental effort-assymetry underlying everyday communication: what is “hard” for a speaker is likely different than what is “hard” for a listener. Zipf described this as follows -- purely from the standpoint of speaker effort, an optimal language $\ell_{speaker}^*$ would tend toward a vocabulary of a single, low-cost word. Given such a language, the full set of potential meanings would be conveyed using only that word, i.e. $\ell_{speaker}^*$ would be fully ambiguous and all possible meanings would need to be disambiguated by a listener. From the standpoint of listener effort, an optimal language $\ell_{listener}^*$ would map all possible meanings to distinct words, removing a listener's need to disambiguate. In this example, speaker effort is related to *production cost* and listener effort to *understanding or disambiguation cost*. Clearly, natural languages fall between the two extremes of $\ell_{speaker}^*$ and $\ell_{listener}^*$. Zipf proposed that the particular lexicon-level properties he observed were a result of these competing forces -- the pressure to jointly minimize speaker and listener effort.\par

@@

In the current work we provide a case-study for the framing presented above. We choose a property of languages that could, in principle, vary freely, but shows strong regularities across languages. The explanadum is why this regularity holds. We examine ambiguity as our property, extending ideas by Piantadosi et al. [-@Piantadosi2011a]. We define a novel measure of efficiency that depends on the interactional behavior of speaker and listener agents. We adopt the reference game as our primary unit of interaction and model language users with the Rational Speech Act (RSA) framework -- a computional model of language use, which is nicely attested by experimental data on interaction. Using these basic ingredients, we show that the property of interest (ambiguity) is prevalent in languages that optimize our measure of efficiency. Further, we show how ambiguity is \textit{used} efficiently during local, in-the-moment interactions. The contributions of this work are twofold -- we derive a novel objective for measuring linguistic efficiency and also show how the reference game framework, in combination with formal models of communication, can be used to connect ideas about system-level efficiencies to in-the-moment language use.\par

@@

### Reference games
Zipf’s example of optimal speaker- and listener-languages took the form of a reference game. We adopt that formulation here, assuming these communication games as our basic unit of analysis. In this framework, speakers and listeners are aware of a set of objects $M$ (*meanings*) and are knowledgeable about the set of possible signals $U$ (*utterances*) that can be used to refer to a given meaning (see Figure 1). Utterances may have different relative costs, operationalized via a prior over utterances $P(U)$. Similarly, meanings differ in the relative degree to which they need to be talked about, operationalized as a prior over meanings $P(M)$\footnote{The prior over meanings are analogous to the \textit{need probabilities} assumed in previous work (Regier, Kemp \& Kay (2015).}. We consider a set of contexts $C$ with an associated prior $P(C)$. Each context $c\in C$ describes a different distribution over meanings e.g. $p(M|C=c_i) \neq p(M|C=c_j)$. Finally, we consider a set of communicative events $e \in E$ where $<u, m, c> = e$ is an utterance-meaning-context triple.\par

@@

### Speakers and listeners
The Rational Speech Act framework (RSA) is a computational-level theory of pragmatic language use, which has been well attested in experimental work examining both speaker and listener communication behavior [@FrankGoodman2012a; @GoodmanFrank2016a]. RSA is a formalization of essential Gricean pragmatic principles -- agents reason about one another and their shared context [@Grice1975a]. We adopt RSA as our representational framework to model Gricean (rational and pragmatic) speaker-listeners in the reference game setting\footnote{See SI section 1 for a more detailed description of basic RSA speaker-listener defintions and to Goodman \& Frank (2016) for an overview of its applications.}.\par

@@

\begin{equation}
\begin{split}
  = \mathbb{E}_{c \sim P(C)}[H_{cross}(P_{speaker}, P_{listener} | c; \ell)]\\
\end{split}
\end{equation}

From an information-theoretic perspective this objective is intuitive: The Cross-Entropy (CE) gives us a measure of dissimilarity between two distributions -- the average number of bits required to communicate under one distribution, given that the “true” distribution differs. In our case, we have an expectation over this term -- the expected difference between the distributions assumed by the speaker $P_{speaker}$ and listener $P_{listener}$ given a set of contexts $C$\footnote{Note that in the single context case $|C|=1$ this objective is simply the speaker-listener Cross-Entropy over utterance-meaning pairs $H_{cross}[P_{speaker}(u, m; \ell), P_{listener}(u, m; \ell)]$}. In other words, an "efficient" language $\ell$ minimizes the distance between what speakers and listeners think.

@@

The task of understanding language is marked by a frequent need to handle various forms of ambiguity: lexical, syntactic, among others [@WasowPerforsBeaver2005a]. The ubiquity of this property, however, has been argued to provide evidence that languages have not been optimized for communication\footnote{Chomsky's (2002) claim that "... language is not well designed, because you have such properties as ambiguity" is analogous to the Zipfian optimal Listener language $\ell_{listener}^*$.} [@Chomsky2002a].

@@

As an example, say we have two objects ($m_1$ and $m_2$), two utterances ($u_1$ and $u_2$), which differ in cost, and two languages ($\ell_1$ and $\ell_2$), which describe different utterance-meaning mappings. In language $\ell_1$, the low-cost $u_1$ can be used to refer to both $m_1$ and $m_2$ ($[[u_1]]_{\ell_1} = \{m_1, m_2 \}$), but the high-cost $u_2$ cannot be used at all ($[[u_2]]_{\ell_1} = \emptyset$). By contrast, in language $\ell_2$, $u_1$ can only refer to $m_1$ and $u_2$ can only refer to $m_2$ ($[[u_1]]_{\ell_2} = \{m_1\}$ and $[[u_2]]_{\ell_2} = \{m_2 \}$). While it is cheaper for a speaker to use $\ell_1$ (because speaking is always low-cost), it is more difficult for a listener (because $u_1$ is ambiguous). Crucially, however, if context is disambiguating then the speaker can use $u_1$ to refer to either $m_1$ or $m_2$ and $\ell_1$ should be preferred to $\ell_2$.\par

In the following experiments we explore two aspects of Piantadosi’s claim. In Simulation 1, we examine the efficient language *structure* aspect of their claim, exploring when the optimal linguistic system $\ell^*$ is most likely to contain ambiguous lexical items. In Simulation 2, we explore an efficient language *use* aspect of the claim -- at what point in a conversation is it useful for a speaker to use ambiguous lexical material?\par

@@

We adopt the hypothesis presented by Piantadosi, formalizing their argument in our current framework -- showing that ambiguity is an efficient property under our CE objective in the reference game setting. We proceed by generating languages with different amounts of contextual support (varying the size of $|C|$). We search the space of languages, examining whether ones which minimize our objective contain ambiguity. If context leads to more efficient communication, then optimal languages under our objective should be more likley to be ambiguous as the amount of context increases.\par

@@

We conduct $N=2000$ simulations. For each simulation we enumerate the set of *valid* languages in which $|U|=|M|=4$ (recall that $U$ is our set of utterances and $M$ our set of meanings). Note that a language $\ell \in L$ is "valid" so long as each possible meaning in $m \in M$ can be referred to by at least one form $u \in U$ (every column of $\ell$ has some non-zero assignment) and each form maps to at least one meaning (every column has some non-zero assignment). For a given simulation, the goal is to find the language $\ell^*$ which minimizes our objective and then check to see if that language contains ambiguity.\par

@@

Following the argument given by Piantadosi et al. [-@Piantadosi2011a], we want to assess the impact of *context* on the presence of ambiguity in optimal languages. To do so we consider four conditions with $n=500$ simulations each (that is, 500 unique sets of $\{P(U), P(M|C), P(C)\}$. Our first is a \textit{one-context} condition ($|C|=1$) -- only a single distribution over meanings $P(M)$. In our \textit{two-context} condition ($|C| = 2$), we consider efficiency under both $P(M|C=c_1)$ as well as $P(M|C=c_2)$. \textit{Three-} and \textit{four-context} conditions corresponding accordingly.\par

@@

In Simulation 1 we explored the degree to which ambiguity is an efficient property of languages when communication is contextualized. Figure 1. plots the proportion of optimal languages under each objective as a function of context size. The red line shows that as the number of contexts increases, so does the probability that an optimal language $\ell^*_{cross}$ contains ambiguity under our CE objective. For comparison we also plot the proportion of speaker-optimal $\ell^*_{speaker}$ (blue line) and listener-optimal $\ell^*_{listener}$ (green line) languages that contain ambiguity. In line with Zipf's predictions, if languages are designed only to minimize speaker effort then optimal languages always contain ambiguity. If languages are designed to minimize listener effort then ambiguity is always avoided.\par

While our results indicate that ambiguity is an efficient property of contextualized language, these simulations assumed that agents could always disambiguate items in context. That is, in our four conditions both agents had perfect knowledge of the relevant conditional distributions ($P(M|C)$). This assumption may be too strong for describing much of day-to-day communication -- we seldom interact with others with perfect knowledge of the current context (or topic) at the start of a conversation. To explore how ambiguity may be \textit{used} efficiently in our framework, we next examine a case in which the listener has imperfect knowledge of context at the start of the conversation, but may infer it from the contents of discourse.\par

@@

We shift focus in Simulation 2 -- instead of examining the landscape of a space of languages under our CE objective, we consider a single language containing ambiguity, examining how that property is used over a discourse. If context is not informative with respect to meaning early in a discourse, but is at later positions, then an efficient speaking strategy would avoid ambiguous material early, saving it for when contextual information is disambiguating. We can consider “context” as analogous to a “topic” of conversation. As an example, imagine a scenario in which a reader is beginning a magazine article. While they may have some knowledge about the article’s topic (perhaps from the title), they may not have complete knowledge of its contents, including the persons or events involved. In this setting, using a low-cost, but ambiguous pronoun early may lead to misunderstanding if context is not informative. But, if by a later position enough contextual information has accumulated, it may be efficient to use the ambiguous material.\par

@@

Over the course of a discourse $D$, the listener tries to infer both the current topic, $c_{current}$, as well as the particular meaning $m$ of a given utterance $u$. That is, we consider updated listener and speakers ($L(m, c|u,D;\ell)$, $S(u|m,c,D;\ell)$) in which both agents can track the history of previous utterances $D$. Importantly, an agent can attempt to infer the current topic of conversation $c_{current}$ using the discourse history $D$.\par

@@

## Speaker agents
We consider three types of speaker models. Our \textit{Full pragmatics} agent, models a speaker who reasons about her listener and also has complete recall of the set of utterances in the discourse $D$. This speaker believes that the listener may not know the current topic $c_{current}$ at the start of the discrouse, but can infer it over the discourse. We compare two baseline models. The first, a \textit{Partial pragmatics} baseline describes a speaker who reasons about a listener, but assumes they have no access to discourse history. The second, a \textit{No pragmatics} speaker who does not consider a listener at all, but produces utterances according to the underlying langauge semantics ($\ell$) and topic probabilities ($p(M|C=c_{current}$) (see SI section 3 for details).\par

@@

Figure 2, (A) shows the empirical probability that a speaker uses an ambiguous utterance as a function of discourse position. The \textit{No pragmatic} baseline uses ambiguous utterances frequently and at a constant rate over the discourse and the \textit{Partial pragmatic} baseline avoids ambiguous utterances entirely. But, the \textit{Full pragmatic} model avoids ambiguous material only at the start of the discourse, employing it increasingly as the discourse proceeds. (C) tracks our CE objective for each model over the discourse. Note that the objective decreases for all three models, primarily driven by the listener updating his belief about the actual topic ($P(C=c_{current}|D)$). However, the objective declines more quickly under the \textit{Full} and \textit{Partial} pragmatic speakers as listener agents are better able to infer the correct context. Additionally, the difference in CE between the \textit{Full}- and \textit{Partial pragmatic} models at the end of the discourse is driven by the reduction in speaker costs described in (B). While speaker effort remains constant in both \textit{No pragmatic} and \textit{Partial pragmatic} baselines, effort declines in the \textit{Full pragmatic} model as she increasingly relies on ambiguous material later in the discourse.\par
